Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2023-04-21 paper/ms.tex
Update abstract

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral disks in the nearby Universe.  
  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies. Previous work ignoring the gas phase of
  galaxies has conclusively demonstrated that bars should slow their rotation
  over time due to their interaction with dark matter halos. We have performed a
  simulation of a Milky Way-like galactic disk hosting a strong bar which
  includes a state-of-the-art model of the interstellar medium and a live dark
  matter halo. In this simulation the bar pattern does not slow down over time,
  and instead remains at a stable, constant rate of rotation. This behavior has
  been observed in previous simulations using more simplified models for the
  interstellar gas, but the apparent lack of secular evolution has remained
  unexplained. We propose that the gas phase of the disk and the dark matter
  halo act in concert to stabilize the bar pattern speed and prevent the bar
  from slowing down or speeding up. We find that in a Milky Way-like disk, a gas
  fraction of only about 5\% is necessary for this mechanism to operate. Our
  result naturally explains why nearly all observed bars rotate rapidly and is
  especially relevant for our understanding of how the Milky Way arrived at its
  present state.

>>>>

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral disks in the nearby Universe.  
  Due to gravitational interactions between the bar and the other components of
  galaxies, it is expected that angular momentum and matter will redistribute
  over long (Gyr) timescales in barred galaxies. Previous work ignoring the gas
  phase of galaxies has conclusively demonstrated that bars should slow their
  rotation over time due to their interaction with dark matter halos. We have
  performed a simulation of a Milky Way-like galactic disk hosting a strong bar
  which includes a state-of-the-art model of the interstellar medium and a live
  dark matter halo. In this simulation the bar pattern does not slow down over
  time, and instead remains at a stable, constant rate of rotation. This
  behavior has been observed in previous simulations using more simplified
  models for the interstellar gas, but the apparent lack of secular evolution
  has remained unexplained. {\bf We find that the presence of the gas phase
  arrests the process by which the dark matter halo slows down a bar, a
  phenomenon we term halo exhaustion. This exhaustion is responsible for
  stabilizing the bar pattern speed.} We find that in a Milky Way-like disk, a
  gas fraction of only about 5\% is necessary for this mechanism to operate. Our
  result naturally explains why nearly all observed bars rotate rapidly and is
  especially relevant for our understanding of how the Milky Way arrived at its
  present state.
